---
title: "Synthetic Retail Sales Script Documentation"
author: "Saravanan Gnanasekaran"
output: html_document
---

**AI Assistance Declaration:** I used ChatGPT for project planning,
synthetic data examples, R learning guidance, and documentation formatting.
Prompts used: "Suggest Markdown formatting and code block examples for
documenting this script in Script_Documentation.Rmd," with my script
description and formatting requirements. The complete prompt and key
response will be recorded in Appendix.md. I reviewed the documentation
against my R Console outputs and displayed chart. Independent Google
Sheets verification and knitting checks are pending. All final calculations
are done by myself. I am responsible for the accuracy and originality of
this work.

# Purpose

The script `analysis.R` examines fictional retail transactions and
summarizes sales by product category. This project demonstrates a basic
R analysis workflow using synthetic data.

# Inputs

The input file, `synthetic_sales.csv`, contains nine fictional transactions
and five columns.

| Column | Description | Imported type |
| --- | --- | --- |
| transaction_id | Fictional transaction identifier | Character |
| category | Electronics, Homeware, or Stationery | Character |
| quantity | Number of items sold | Integer |
| unit_price | Price per item | Integer |
| sales_amount | Transaction sales amount supplied in the CSV | Integer |

No currency is specified. The dataset contains no real customer information.

The intended transaction calculation is quantity multiplied by unit price.
The script summarizes the supplied sales amounts rather than recalculating
each transaction.

# Requirements

- R, accessed through RStudio in Posit Cloud or a local installation.
- No additional packages are required for `analysis.R`.
- The CSV and script must be accessible from the project working directory.
- Rendering this document requires R Markdown tooling, which is separate
  from the analysis script's requirements.

The R version used has not yet been recorded.

# Script Workflow

1. Import the CSV into an object named `sales`.
2. Preview the records using `head()`.
3. Inspect column types using `str()`.
4. Check dataset dimensions using `dim()`.
5. Check for missing values using `anyNA()`.
6. Summarize sales by category using `aggregate()`.
7. Display category totals using `barplot()`.

# Code Examples

These examples reproduce code already used in the project.
They are displayed without execution because `eval=FALSE`.

## Import the Dataset

```{r import-example, eval=FALSE}
sales <- read.csv("synthetic_sales.csv")
```

The relative path assumes that the CSV is in the working directory.
The imported data is stored in `sales`.

## Inspect the Dataset

```{r inspection-example, eval=FALSE}
head(sales)
str(sales)
dim(sales)
```

These instructions preview records, identify column types, and show
the number of rows and columns.

## Check Missing Values

```{r missing-values-example, eval=FALSE}
anyNA(sales)
```

A result of `FALSE` means no missing values were detected.
It does not establish that all values are accurate.

## Summarize Sales by Category

```{r category-summary-example, eval=FALSE}
category_sales <- aggregate(sales_amount ~ category,
                            data = sales,
                            FUN = sum)
category_sales
```

This instruction groups transactions by category and adds their supplied
sales amounts.

## Display the Chart

```{r chart-example, eval=FALSE}
# Replace this comment with the exact barplot() instruction
# from your saved analysis.R script.
```

The chart was displayed successfully, but the exact chart instruction
has not been supplied for this documentation draft.

# Outputs

## Reported R Results

The imported dataset contained nine rows and five columns.
The missing-value check returned `FALSE`.

The following table reproduces the reported R Console output.
Independent Google Sheets verification is pending.

| Category | Sales amount |
| --- | ---: |
| Electronics | 85 |
| Homeware | 56 |
| Stationery | 40 |

## Chart

The script displays a bar chart in the Plots panel with:

> barplot(category_sales$sales_amount,
+         names.arg = category_sales$category,
+         main = "Synthetic Sales by Category",
+         xlab = "Product Category",
+         ylab = "Sales Amount",
+         col = "steelblue")

The script does not save tables or chart files.
Knitting this document produces a separate documentation output.

# Verification

| Check | Observed result or status |
| --- | --- |
| CSV import | The sales object appeared in the Environment |
| Column types | Two character columns and three integer columns |
| Dataset dimensions | Nine rows and five columns |
| Missing values | anyNA(sales) returned FALSE |
| Chart appearance | Visible bars were consistent with reported category totals |
| Spreadsheet transaction calculations | Pending confirmation |
| Independent category totals in Google Sheets | Pending |
| Exact chart code inserted into this document | Pending |
| Document knitted and visually inspected | Pending |

The results above were reported from the student's R session.
Visual consistency is not an independent calculation check.

# Limitations

- Nine synthetic records cannot establish real retail trends.
- No currency is specified.
- The script does not currently check duplicate identifiers, invalid
  categories, negative values, or transaction arithmetic consistency.
- The project provides descriptive totals, not forecasts or statistical
  inference.
- The displayed code chunks do not execute during knitting.
- Independent spreadsheet verification remains pending.

# AI Assistance Disclosure

ChatGPT assisted with planning, synthetic data examples, R explanations,
instructional code patterns, and documentation drafting.

Date of assistance: October 7, 2026 (America/Toronto).

Model/version: Not recorded; add the displayed model if available.

The student must review this draft against the saved script, replace the
chart placeholder, and record subsequent prompts and checks in Appendix.md.